\section{Miscellaneous Routines}
\label{s:misc}

\begin{fctbox}
\textcolor{output}{\textit{ier}} = cgio\_flush\_to\_disk(\textcolor{input}{int cgio\_num}); & - w m \\
\textcolor{output}{\textit{ier}} = cgio\_library\_version(\textcolor{input}{int cgio\_num}, \textcolor{output}{\textit{char *version}}); & r w m \\
\textcolor{output}{\textit{ier}} = cgio\_file\_version(\textcolor{input}{int cgio\_num}, \textcolor{output}{\textit{char *file\_version}}, & r w m \\
~~~~~~~\textcolor{output}{\textit{char *creation\_date}}, \textcolor{output}{\textit{char *modified\_date}}); & \\
\hline
call cgio\_flush\_to\_disk\_f(\textcolor{input}{cgio\_num}, \textcolor{output}{\textit{ier}}) & - w m \\
call cgio\_library\_version\_f(\textcolor{input}{cgio\_num}, \textcolor{output}{\textit{version}}, \textcolor{output}{\textit{ier}}) & r w m \\
call cgio\_file\_version\_f(\textcolor{input}{cgio\_num}, \textcolor{output}{\textit{file\_version}}, \textcolor{output}{\textit{creation\_date}}, & r w m \\
~~~~~\textcolor{output}{\textit{modified\_date}}, \textcolor{output}{\textit{ier}}) & \\
\end{fctbox}

\noindent
\textbf{\textcolor{input}{Input}/\textcolor{output}{\textit{Output}}}

\begin{Ventryi}{\texttt{modified\_date}}\raggedright
\item [\texttt{cgio\_num}]
      Database identifier.
\item [\texttt{version}]
      32-byte character string containing the database library version.
\item [\texttt{file\_version}]
      32-byte character string containing the database file version.
\item [\texttt{creation\_date}]
      32-byte character string containing the database file creation date.
\item [\texttt{modified\_date}]
      32-byte character string containing the last modification date
      for the database file.
\item [\texttt{ier}]
      Error status.
\end{Ventryi}

\subsection{Function Descriptions}

\subsubsection{\texttt{cgio\_flush\_to\_disk}} \label{flush_to_disk}
    \noindent
    Forces any buffered data in the database manager to be written
    to disk. Returns 0 if successfull, else an error code.

\subsubsection{\texttt{cgio\_library\_version}} \label{library_version}
    \noindent
    Gets the current library version for the database given by \texttt{cgio\_num}.
    The version is returned in \texttt{version} which is of maximum length
    \texttt{CGIO\_MAX\_VERSION\_LENGTH} (32). In C, \texttt{version} should
    be dimensioned at least 33 in the calling routine to allow for the
    terminating '0'.
    The function returns 0 if successfull, else an error code.

\subsubsection{\texttt{cgio\_file\_version}} \label{file_version}
    \noindent
    Gets the version, creation and last modified dates, for the database
    file given by \texttt{cgio\_num}. The version is returned in
    \texttt{file\_version}, which is of maximum length
    \texttt{CGIO\_MAX\_VERSION\_LENGTH} (32). The creation date is returned in
    \texttt{creation\_date}, and the last modified date in \texttt{modified\_date},
    which are of maximum length \texttt{CGIO\_MAX\_DATE\_LENGTH} (32).
    In C, these should be dimensioned at least 33 in the calling routine
    to allow for the terminating '0'.
    The function returns 0 if successfull, else an error code.

