\section{Applications Software}
\label{s:applications}
\thispagestyle{plain}

The development of CGNS-compliant applications, e.g., grid generators,
postprocessors, and the like, has not been a direct undertaking of the
CGNS team.
Rather, it has been the intent to make the attractiveness of
interoperable CFD applications, together with general acceptance of
the CGNS standard by Boeing, NASA, and others, sufficient to induce
applications developers to incorporate CGNS I/O into their offerings.

Several CGNS-compatible applications have indeed been developed, and
more continue to appear.
This section lists the known applications at the time of publication.
Please contact Tony Iannetti (\textit{Anthony.C.Iannetti@nasa.gov})
to modify or add to the lists.

\subsection{CGNS-Compatible Applications}

This section lists major software packages that read and/or write
CGNS files.
The software listed here is taken from Table 1 of AIAA Paper 2002-0752,
and is for informational purposes only.
The CGNS Project Group cannot endorse specific software packages.

\setlength{\LTleft}{\fill}
\setlength{\LTright}{\fill}
\setlength{\LTcapwidth}{5.0in}
\settowidth{\tmplength}{Advanced Visual Sys.}
\settowidth{\tmplengtha}{ANSYS ICEM CFD}
\setlength{\Pwidth}{\linewidth-4\tabcolsep-\tmplength-\tmplengtha}
\begin{longtable}{@{}l l >{\itshape}p{\Pwidth}@{}}
\caption[CGNS-Compatible Grid Generators and Pre-Processors]{\textbf{CGNS-Compatible Grid Generators and Pre-Processors}}
\label{t:gridgen}
\\ \hline\hline \\*[-2ex]
\bold{Organization} & \bold{Application} & \bold{Contact}
\\*[1ex] \hline\hline \\*[-2ex]
\endfirsthead

\multicolumn{3}{l}{{\bfseries\autoref{t:gridgen}: CGNS-Compatible Grid Generators} (\emph{Continued})}
\\*[1ex] \hline\hline \\*[-2ex]
\bold{Organization} & \bold{Application} & \bold{Contact}
\\*[1ex] \hline\hline \\*[-2ex]
\endhead

\\*[-2ex]\hline
\multicolumn{3}{r}{\emph{Continued on next page}} \\
\endfoot
\\*[-2ex] \hline\hline
\endlastfoot
\textit{Commercial} \\
Advanced Visual Sys. & ANSYS ICEM CFD & \url{http://cfl3d.larc.nasa.gov/Cfl3dv6/cfl3dv6.html#contact} \kill
Ansys, Inc.     & ANSYS ICEM CFD & \url{ansysinfo@ansys.com} \\
NUMECA          & IGG           & \url{http://www.numeca.be/index.php?id=offices} \\
Pointwise, Inc. & GRIDGEN       & \url{http://www.pointwise.com/pointwise/contact.shtml} \\
\\
\textit{Industry} \\
Boeing Rocketdyne        & APPT   & \\
\end{longtable}

\enlargethispage{2\baselineskip}

\setlength{\LTleft}{\fill}
\setlength{\LTright}{\fill}
\begin{longtable}{@{}l l >{\itshape}p{\Pwidth}@{}}
\caption[CGNS-Compatible Flow Solvers]{\textbf{CGNS-Compatible Flow Solvers}}
\label{t:solvers}
\\ \hline\hline \\*[-2ex]
\bold{Organization} & \bold{Application} & \bold{Contact}
\\*[1ex] \hline\hline \\*[-2ex]
\endfirsthead

\multicolumn{3}{l}{{\bfseries\autoref{t:solvers}: CGNS-Compatible Flow Solvers} (\emph{Continued})}
\\*[1ex] \hline\hline \\*[-2ex]
\bold{Organization} & \bold{Application} & \bold{Contact}
\\*[1ex] \hline\hline \\*[-2ex]
\endhead

\\*[-2ex]\hline
\multicolumn{3}{r}{\emph{Continued on next page}} \\
\endfoot
\\*[-2ex] \hline\hline
\endlastfoot
\textit{Commercial} \\
Advanced Visual Sys. & ANSYS ICEM CFD & \url{http://cfl3d.larc.nasa.gov/Cfl3dv6/cfl3dv6.html#contact} \kill
Ansys, Inc.    & CFX5         & \url{ansysinfo@ansys.com} \\
               & CFX-TASCFlow & \url{ansysinfo@ansys.com} \\
Aerosoft, Inc. & GASP         & \url{http://www.aerosft.com/Misc/contact.php} \\
CD ADAPCO      & STAR-CD      & \url{http://www.cd-adapco.com/about/global_locations.html} \\
Fluent, Inc.   & Fluent       & \url{http://www.fluent.com/contact/} \\
Newmerical Tech. & FENSAP     & \url{http://www.newmerical.com/public/eng/index.php?pageId=about&language=en} \\
NUMECA         & EURANUS      & \url{http://www.numeca.be/index.php?id=offices} \\
\newpage
\textit{Government} \\
DLR/MTU        & TRACE    & \url{Frank.Eulitz@dlr.de} \\
NASA Langley   & PEGSUS   & \url{Pieter.G.Buning@nasa.gov} \\
               & OVERFLOW & \url{Pieter.G.Buning@nasa.gov} \\
               & Cart3D   & \url{http://people.nas.nasa.gov/~aftosmis/} \\
NASA Glenn     & NCC      & \url{Nan-Suey.Liu-1@nasa.gov} \\
NASA Langley   & CFL3D    & \url{http://cfl3d.larc.nasa.gov/Cfl3dv6/cfl3dv6.html#contact} \\
               & UPS      & \\
NPARC Alliance & WIND     & \url{npsupport@info.arnold.af.mil} \\
ONERA          & elsA     & \url{http://elsa.onera.fr/elsA/contacts.html} \\
\\
\textit{Industry} \\
Boeing Seattle           & TLNS3D & \\
Rolls-Royce Allison      & ADPAC  & \\
Rolls-Royce Oxford       & HYDRA  & \url{pierre.moinier@comlab.ox.ac.uk}
\end{longtable}

\setlength{\LTleft}{\fill}
\setlength{\LTright}{\fill}
\begin{longtable}{@{}l l >{\itshape}p{\Pwidth}@{}}
\caption[CGNS-Compatible Post Processors]{\textbf{CGNS-Compatible Post Processors}}
\label{t:postprocess}
\\ \hline\hline \\*[-2ex]
\bold{Organization} & \bold{Application} & \bold{Contact}
\\*[1ex] \hline\hline \\*[-2ex]
\endfirsthead

\multicolumn{3}{l}{{\bfseries\autoref{t:postprocess}: CGNS-Compatible Post Processors} (\emph{Continued})}
\\*[1ex] \hline\hline \\*[-2ex]
\bold{Organization} & \bold{Application} & \bold{Contact}
\\*[1ex] \hline\hline \\*[-2ex]
\endhead

\\*[-2ex]\hline
\multicolumn{3}{r}{\emph{Continued on next page}} \\
\endfoot
\\*[-2ex] \hline\hline
\endlastfoot
\textit{Commercial} \\
Advanced Visual Sys. & ANSYS ICEM CFD & \url{http://cfl3d.larc.nasa.gov/Cfl3dv6/cfl3dv6.html#contact} \kill
Advanced Visual Sys. & AVS/Express  & \url{http://www.avs.com/contact/} \\
Aerosoft, Inc.       & GASP         & \url{http://www.aerosft.com/Misc/contact.php} \\
Tecplot, Inc.        & Tecplot      & \url{http://www.tecplot.com/contact_us/} \\
CEI                  & Ensight      & \url{http://www.ensight.com/contact-us-3.html} \\
                     & Ensight Gold & \url{http://www.ensight.com/contact-us-3.html} \\
Fluent, Inc.         & Fluent       & \url{http://www.fluent.com/contact/} \\
Ansys, Inc.          & ANSYS ICEM CFD & \url{ansysinfo@ansys.com} \\
Intelligent Light    & Fieldview    & \url{fieldview@ilight.com} \\
Newmerical Tech.     & DROP3D       & \url{http://www.newmerical.com/public/eng/index.php?pageId=about&language=en} \\
                     & ICE3D        & \url{http://www.newmerical.com/public/eng/index.php?pageId=about&language=en} \\
NUMECA               & CFView       & \url{http://www.numeca.be/index.php?id=offices} \\
\\
\textit{Government} \\
NASA Ames            & PLOT3D       & \url{Cetin.C.Kiris@nasa.gov}
\end{longtable}

\newpage
\subsection{CGNS Utilities}

A variety of CGNS utility packages have been developed, both by CGNS
team members and by CGNS users.
While not formally a part of CGNS itself, these utilities are available
from SourceForge, at \textit{http://sourceforge.net/projects/cgns},
and/or from the CGNS web site, at \textit{http://cgns.sourceforge.net/}.

\begin{Ventryi}{Plot3dg\_to\_CGNS}
\item [Plot3dg\_to\_CGNS]
      Converts a ``standard'' Plot3D grid file to a CGNS grid file.\\
      Contact: Chris Rumsey (\textit{Christopher.L.Rumsey@nasa.gov})
\item [CGNS\_to\_Plot3d]
      Converts a CGNS file with a cell-centered solution to ``standard''
      Plot3D grid and solution files.\\
      Contact: Chris Rumsey (\textit{Christopher.L.Rumsey@nasa.gov})
\item [CGNS\_readhist]
      Reads a CGNS file and writes history data to a formatted file.\\
      Contact: Chris Rumsey (\textit{Christopher.L.Rumsey@nasa.gov})
\item [FTU]
      Converts CGNS files to and from Plot3D format.
      Has a text-based menu allowing the manipulation of a CGNS base.\\
      Contact: Bob Bush (\textit{bushrh@pweh.com})
\item [ADFviewer]
      An ADF/CGNS file viewer and editor, written in \textit{Tcl/Tk}.
      The GUI allows complete editing of ADF/CGNS files, and handles
      links and data conversions.
      It includes built-in CGNS node and label information, and
      documentation support through HTML (including access to the ADF
      and CGNS docs).\\
      Contact: Bruce Wedan (\textit{brucewedan@hotmail.com})
\item [CGNS Tools]
      A variety of tools for viewing, editing, and manipulating the
      contents of CGNS files; requires \textit{Tcl/Tk} 8.3.
      Included are:
      \begin{itemize*}
      \item \textit{CGNSplot}, for plotting grids
      \item \textit{CGNScalc}, a calculator using data in CGNS files
      \item Utilities for converting from Patran, and from/to PLOT3D and
            Tecplot
      \item Utilities for modifying the grid location
            (vertex or cell-center), solution variables
            (primitive or conservative), and data class
            (\fort{Dimensional}, \fort{NormalizedByDimensional}, or
            \fort{NormalizedByUnknownDimensional})
      \item Utilities for creating a subset or interpolated version of
            a CGNS file
      \end{itemize*}
      These may be used as stand-alone utilities, or initiated from
      \textit{ADFviewer}, which is included.\\
      Contact: Bruce Wedan (\textit{brucewedan@hotmail.com})
\item [CGNS Viewer]
      An alpha release of another ADF/CGNS file viewer and editor.
      Requires the GTK+ library version 1.2 or newer (installed by default
      on newer Linux systems using Gnome; otherwise found at
      {\itshape\url{http://www.gtk.org/}}).\\
      Contact: Christian Lundh (\textit{xistan@yahoo.se})
\item [PyCGNS]
      A Python binding to the ADF Core and Mid-Level Library.
      For more information, see the pyCGNS home page at
      {\itshape\url{http://elsa.onera.fr/CGNS/releases/}}.\\
      Contact: Marc Poinot (\textit{Marc.Poinot@onera.fr})
\item [ADFM]
      A development version of software for managing an in-memory
      representation of ADF trees.
      For more information, see ``Proposal and Prototype for an in-memory
      representation of ADF trees'' at
      {\itshape\url{http://cgns.sourceforge.net/utilities/ADFM/ADFM.pdf}}.\\
      Contact: Marc Poinot (\textit{Marc.Poinot@onera.fr})
\item [CGNS++]
      A C++ binding to the ADF Core and Mid-Level Library.
      For more information, see the CGNS++ home page at
      \textit{http://cgnspp.sourceforge.net/}.\\
      Contact: Manuel Kessler (\textit{kessler@iag.uni-stuttgart.de})
\end{Ventryi}
