\section{History of ADF Version Releases}
\label{s:history}
\thispagestyle{plain}

This section contains a list of all the versions of ADF Software
Libraries that have been released since the first release in December
1995.
Along with each release, we include a brief synopsis of the improvements
made in this version.
In the future, newer versions may be implemented and released and they
will also be accessible on the World Wide Web.

\begin{Ventryi}{ADF AXX (Future Release)}
\item [ADF A01 (December 1995)]
      This was the first release of the software.
      It includes the majority of the subroutines that are summarized
      in this document, and it was internally distributed for use at
      Boeing and to members of the working group (NASA and McDonnell
      Douglas).
\item [ADF A02, A03 (July 1996)]
      This release includes improvements and some bug fixes for the A01
      version.
\item [ADF B01 (September 1996)]
      This release includes the first release of the ADF library and
      with the following subroutines added:
      \hyperlink{sub:Delete}{Delete a Node},
      \hyperlink{sub:Put_Name}{Add a Name (Change) on a Node},
      \hyperlink{sub:Move_Child}{Change the Parent Node (move a Child)}.
\item [ADF C00 (October 1996)]
      This release includes the performance enhancements and support
      for more platforms (Cray T90, SGI 6.2, DEC Alpha, Intel Paragon).
\item [ADF AXX (Future Release)]
      It is anticipated that there will be future releases of the ADF
      core libraries, and they will include at a minimum the following
      additional routines (plus others that will provide enhancements
      as they are needed):
      Delete an Existing Database and
      \hyperlink{sub:Database_Garbage_Collection}{Garbage Collection}.
      Garbage Collection will redistribute the ADF file to use
      free-space that is not located at the end of the file.
\end{Ventryi}
