\section{File Operations}
\label{s:fileops}
\thispagestyle{plain}

\subsection{Opening and Closing a CGNS File}
\label{s:openclose}

\begin{fctbox}
\textcolor{output}{\textit{ier}} = cg\_open(\textcolor{input}{char *filename}, \textcolor{input}{int mode}, \textcolor{output}{\textit{int *fn}}); & r w m \\
\textcolor{output}{\textit{ier}} = cg\_version(\textcolor{input}{int fn}, \textcolor{output}{\textit{float *version}});         & r w m \\
\textcolor{output}{\textit{ier}} = cg\_close(\textcolor{input}{int fn});                           & r w m \\
\textcolor{output}{\textit{ier}} = cg\_is\_cgns(\textcolor{input}{const char *filename});                           & r w m \\
\hline
call cg\_open\_f(\textcolor{input}{filename}, \textcolor{input}{mode}, \textcolor{output}{\textit{fn}}, \textcolor{output}{\textit{ier}})          & r w m \\
call cg\_version\_f(\textcolor{input}{fn}, \textcolor{output}{\textit{version}}, \textcolor{output}{\textit{ier}})              & r w m \\
call cg\_close\_f(\textcolor{input}{fn}, \textcolor{output}{\textit{ier}})                         & r w m \\
call cg\_is\_cgns\_f(\textcolor{input}{filename}, \textcolor{output}{\textit{ier}})                         & r w m \\
\end{fctbox}

\noindent
\textbf{\textcolor{input}{Input}/\textcolor{output}{\textit{Output}}}

\begin{Ventryi}{\texttt{filename}}\raggedright
\item [\texttt{filename}]
      Name of the CGNS file, including path name if necessary.
      There is no limit on the length of this character variable.
      (\textcolor{input}{Input})
\item [\texttt{mode}]
      Mode used for opening the file.
      The modes currently supported are \texttt{CG\_MODE\_READ},
      \texttt{CG\_MODE\_WRITE}, and \texttt{CG\_MODE\_MODIFY}.
      For backwards compatibility, the equivalent mode
      names \texttt{MODE\_READ}, \texttt{MODE\_WRITE}, and
      \texttt{MODE\_MODIFY} are still supported, but it is recommended
      that the new names be used.
      (\textcolor{input}{Input})
\item [\texttt{fn}]
      CGNS file index number.
      (\textcolor{input}{Input} for \texttt{cg\_version} and
      \texttt{cg\_close}; \textcolor{output}{\textit{output}} for
      \texttt{cg\_open})
\item [\texttt{version}]
      CGNS version number.
      (\textcolor{output}{\textit{Output}})
\item [\texttt{ier}]
      Error status.
      (\textcolor{output}{\textit{Output}})
\end{Ventryi}

The function \texttt{cg\_open} must always be the first one called.
It opens a CGNS file for reading and/or writing and returns an index
number \texttt{fn}.
The index number serves to identify the CGNS file in subsequent
function calls.
Several CGNS files can be opened simultaneously.
The current limit on the number of files opened at once depends on the
platform.
On an SGI workstation, this limit is set at 100 (parameter
\texttt{FOPEN\_MAX} in \textit{stdio.h}).

The file can be opened in one of the following modes:

\begin{Ventryic}{\texttt{CG\_MODE\_MODIFY}}
\item [\texttt{CG\_MODE\_READ}]
      Read only mode.
\item [\texttt{CG\_MODE\_WRITE}]
      Write only mode.
\item [\texttt{CG\_MODE\_MODIFY}]
      Reading and/or writing is allowed.
\end{Ventryic}

When the file is opened, if no \texttt{CGNSLibraryVersion\_t} node is
found, a default value of 1.05 is assumed for the CGNS version number.
Note that this corresponds to an old version of the CGNS standard, that
doesn't include many data structures supported by the current standard.

The function \texttt{cg\_close} must always be the last one called.
It closes the CGNS file designated by the index number \texttt{fn} and
frees the memory where the CGNS data was kept.
When a file is opened for writing, \texttt{cg\_close} writes all the
CGNS data in memory onto disk prior to closing the file.
Consequently, if is omitted, the CGNS file is not written properly.

In order to reduce memory usage and improve execution speed,
large arrays such as grid coordinates or flow solutions are not actually
stored in memory.
Instead, only their ADF (or HDF) ID numbers are kept in memory for future
reference.
When a CGNS file is open in writing mode, large arrays passed to the
library are immediately written into the CGNS file, directly under the
root node.
When the file is closed, these arrays are moved to their appropriate
location in the CGNS tree.

\subsection{Configuring CGNS Internals}
\label{s:configure}

\begin{fctbox}
\textcolor{output}{\textit{ier}} = cg\_configure(\textcolor{input}{int option}, \textcolor{input}{void *value}); & r w m \\
\end{fctbox}

\noindent
\textbf{\textcolor{input}{Input}/\textcolor{output}{\textit{Output}}}

\begin{Ventryi}{\texttt{option}}\raggedright
\item [\texttt{option}]
      The option to configure, currently one of
      \texttt{CG\_CONFIG\_ERROR}, \texttt{CG\_CONFIG\_COMPRESS},
      \texttt{CG\_CONFIG\_SET\_PATH}, or \texttt{CG\_CONFIG\_ADD\_PATH}
      as defined in \textit{cgnslib.h}.
      (\textcolor{input}{Input})
\item [\texttt{value}]
      The value to set, type cast as \texttt{void *}.
      (\textcolor{input}{Input})
\item [\texttt{ier}]
      Error status.
      (\textcolor{output}{\textit{Output}})
\end{Ventryi}

The function \texttt{cg\_configure} allows certain CGNS library internal
options to be configured.
The currently supported options and expected values are:

\begin{Ventryi}{\texttt{CG\_CONFIG\_SET\_PATH}}
\item [\texttt{CG\_CONFIG\_ERROR}]
      This allows an error call-back function to be defined by the
      user.
      The value should be a pointer to a function to receive the error.
      The function is defined as \texttt{void err\_callback(int is\_error,
      char *errmsg)}, and will be called for errors and warnings.
      The first argument, \texttt{is\_error}, will be 0 for warning
      messages, 1 for error messages, and $-1$ if the program is
      going to terminate (i.e., a call to \texttt{cg\_error\_exit()}).
      The second argument is the error or warning message.
      If this is defined, warning and error messages will go to the
      function, rather than the terminal.
      A \texttt{value} of \texttt{NULL} will remove the call-back
      function.
\item [\texttt{CG\_CONFIG\_COMPRESS}]
      When a CGNS file is closed after being opened in modify mode, the
      normal operation of the CGNS library is to rewrite the file if
      there is unused space.
      This happens when nodes have been rewritten or deleted.
      Setting \texttt{value} to 0 will prevent the library from
      rewriting the file, and setting it to 1 will force the rewrite.
      The default value is $-1$.
\item [\texttt{CG\_CONFIG\_SET\_PATH}]
      Sets the search path for locating linked-to files.
      The argument \texttt{value} should be a character string
      containing one or more directories, formatted the same as for the
      \texttt{PATH} environment variable.
      This will replace any current settings.
      Setting \texttt{value} to \texttt{NULL} will remove all paths.
\item [\texttt{CG\_CONFIG\_ADD\_PATH}]
      Adds a directory, or list of directories, to the linked-to file
      search path.
      This is the same as \texttt{CG\_CONFIG\_SET\_PATH}, but adds to
      the path instead of replacing it.
\end{Ventryi}

There is no Fortran counterpart to function \texttt{cg\_configure}.
