\pdfbookmark[1]{Overview}{overview}
\addtocontents{toc}{\protect\contentsline {section}{Overview}{\thepage}{overview}}
\hypertarget{overview}{}
\section*{Overview}
\thispagestyle{plain}

This document contains the Standard Interface Data Structures (SIDS)
definitions for the CFD General Notation System (CGNS) project; this
project was originally a NASA-funded contract under the AST program, but
control has now been completely transferred to a public forum known as
the CGNS Steering Committee.

The purpose of this document is to scope the information that should be
communicated between various CFD application codes; the target is 3--D
multizone, compressible Navier-Stokes analysis.
Attention in this document is not focussed on I/O routines or formats,
but on the precise description of data that should be present in the I/O
of a CFD code or in a CFD database.

This document therefore contains a precise definition of information
pertinent to a CGNS database.
Specifically, the following information is addressed:
\begin{itemize}
\item grid coordinates and elements
\item flow solution data, including nondimensional parameters
\item multizone interface connectivity, including abutting and overset
\item boundary conditions
\item flow equation descriptions
\item time-dependent flow
\item reference states
\item dimensional units and nondimensionalization information associated
      with data
\item convergence history
\item association to geometry definition
\item topologically based hierarchical structures
\end{itemize}
This information is encoded into C-like data structures.

This version of the ``Standard Interface Data Structures'' represents a
draft revision of the AIAA Recommended Practice R-101A-2005.
This document has not undergone a formal approval vote by the standards
committee and is therefore not an approved AIAA consensus document.
At the time of release of the next major version of the CGNS data
structures, this draft document will be voted on by the committee
members and, pending approval, become the next iteration of the AIAA
Recommended Practice.
To obtain bound paper copies of the current approved version of this
document (corresponding to SIDS Document Version 2.4.5, CGNS Version
2.4), please visit the
\href{http://www.aiaa.org/content.cfm?pageid=363&id=1657}{AIAA
Online Store}.
The current Recommended Practice is also available as a
\href{http://www.grc.nasa.gov/www/cgns/sids/aiaa/R_101A_2005.pdf}{PDF file}
(1.01M, 200 pages).

Questions and comments on this document are welcome and should be
directed to one of the following:
\begin{center}
\begin{tabular}{l@{\qquad\qquad}l}
Charlie Towne               & Diane Poirier \\
MS 86-7                     & ICEM CFD Engineering \\
NASA Glenn Research Center  & 2855 Telegraph Ave Suite 501 \\
Cleveland, OH \ 44135-3191  & Berkeley, CA  94705 \\
(216) 433-5851              & (510) 549-1890 \\
(216) 977-7500 (FAX)        & (510) 841-8523 (FAX) \\
e-mail: towne@nasa.gov      & e-mail: Diane.Poirier@ansys.com
\end{tabular}
\end{center}

